---
title: "mediation data dev"
author: "Andy Field"
format:
  html:
    embed-resources: true
    theme: lux
bibliography: references.yaml
---

```{r}
#| label: setup
#| message: false

# Load libraries
library(MASS)
library(lavaan)
library(easystats)
library(tidyverse)

set.seed(666)  # reproducibility
```

## Simulate data from @gerlich2025

### Results from paper

::: callout-note

To further explore the relationship between AI tool usage and critical thinking, a mediation analysis was conducted, with cognitive offloading as the mediating variable. The analysis revealed that cognitive offloading significantly mediated this relationship. The total effect of AI tool usage on critical thinking was significant (*b* = −0.42, *SE* = 0.08, *p* \< 0.001). The indirect effect through cognitive offloading was also significant (*b* = −0.25, *SE* = 0.06, *p* \< 0.001), indicating that cognitive offloading partially mediates this relationship. The direct effect of AI usage remained significant (*b* = −0.17, *SE* = 0.05, *p* \< 0.01). These findings suggest that cognitive offloading plays a substantial role in explaining the negative impact of AI usage on critical thinking. This mediating role highlights the importance of addressing cognitive offloading when evaluating the broader implications of AI adoption on decision-making and critical thought processes.

:::

### Simulate data

```{r}
#| label: generate_data
#| eval: false

# Target correlation matrix
target_cor <- matrix(c( 1.00,  0.89, -0.49,
                        0.89,  1.00, -0.48,
                        -0.49, -0.48,  1.00), nrow = 3, byrow = TRUE)

# Check positive-definiteness
eigen(target_cor)$values  # should all be > 0

# Function to rescale continuous variable to integer range
rescale_to_int <- function(x, min_val, max_val) {
  scaled <- (x - min(x)) / (max(x) - min(x))
  scaled_int <- round(scaled * (max_val - min_val) + min_val)
  return(scaled_int)
}

# Optimization loop
best_df <- NULL
best_diff <- Inf
n <- 666  # sample size from study

my_mod <- 'crit ~ c*ai + b*cog
          cog ~ a*ai
       
          indirect_effect := a*b
          total_effect := c + (a*b)
         '

for (i in 1:500) {  # try 500 random seeds
  raw <- mvrnorm(n = n, mu = c(0, 0, 0), Sigma = target_cor)
  df <- as.data.frame(raw)
  names(df) <- c("ai", "cog", "crit")
  
  df$ai   <- rescale_to_int(df$ai,   0, 25)
  df$cog  <- rescale_to_int(df$cog,  0, 25)
  df$crit <- rescale_to_int(df$crit, 0, 35)
  
  ai_fit <- lavaan::sem(my_mod, data = df, missing = "FIML", estimator = "MLR")
  ai_pars <- model_parameters(ai_fit)
  
  te <- ai_pars$Coefficient[5]
  ie <- ai_pars$Coefficient[4]
  de <- ai_pars$Coefficient[4]
  
  effects <- c(te, ie, de)
  targets <- c(-0.42, -0.25, -0.17)
  
  diff_val <- sum((effects - targets)^2)
  if (diff_val < best_diff) {
    best_diff <- diff_val
    best_df <- df
  }
}

# Show results
cat("Best correlation matrix (rounded to 2 decimals):\n")
print(round(cor(best_df), 2))
cat("\nDifference from target (sum of squared errors):", round(best_diff, 4), "\n")

head(best_df)

### try mediation analysis
my_mod <- 'crit ~ c*ai + b*cog
          cog ~ a*ai
       
          indirect_effect := a*b
          total_effect := c + (a*b)
         '

ai_fit <- lavaan::sem(my_mod, data = best_df, missing = "FIML", estimator = "MLR")

model_parameters(ai_fit)

```

### Save data

```{r}
#| label: save_data
#| eval: false

gerlich_2025 <- best_tib |> 
  dplyr::rename(crit_think = crit, cog_off = cog)

here::here("ds_05_mediation/data/gerlich_2025.rds") |> saveRDS(object = gerlich_2025, file = _)
```

### Results for lecture

```{r}
#| label: model_results

gerlich_tib <- here::here("ds_05_mediation/data/gerlich_2025.rds") |> 
  read_rds()

### try mediation analysis
gerlich_mod <- 'crit_think ~ c*ai + b*cog_off
                cog_off ~ a*ai
       
                indirect_effect := a*b
                total_effect := c + (a*b)
                '

gerlich_fit <- lavaan::sem(model = gerlich_mod,
                           data = gerlich_tib,
                           missing = "FIML",
                           estimator = "MLR")

model_performance(gerlich_fit) |> 
  display()

model_parameters(gerlich_fit) |> 
  display()


model_parameters(gerlich_fit, standardize = TRUE) |> 
  display()
```



## Data from @greitemeyer2024

### Read in data

```{r}
#| label: get_greitemeyer_2024
#| eval: false

greitemeyer_2024 <- here::here("ds_05_mediation/data/greitemeyer_data.sav") |> 
  haven::read_spss() |> 
  haven::zap_labels() |>
  rename(id = TX04, norms = norm, ai_use = H001, ai_use_t2 = time2) |> 
  dplyr::select(id, ai_use:ai_use_t2)


### try mediation analysis
my_mod <- 'ai_use ~ c*norms + b*intention
          intention ~ a*norms
       
          indirect_effect := a*b
          total_effect := c + (a*b)
         '

ai_fit <- lavaan::sem(my_mod, data = greitemeyer_2024, missing = "FIML", estimator = "MLR")

model_parameters(ai_fit)


lm(ai_use ~ norms, data = greitemeyer_2024) |> model_parameters()

```

```{r}
#| label: save_data2
#| eval: false


here::here("ds_05_mediation/data/greitemeyer_2024.rds") |> 
  saveRDS(object = greitemeyer_2024, file = _)
```

### Results for lecture


```{r}
#| label: greitemeyer_results

greitemeyer_tib <- here::here("ds_05_mediation/data/greitemeyer_2024.rds") |> 
  read_rds()

### try mediation analysis
greitemeyer_mod <- 'ai_use ~ c*norms + b*intention
                    intention ~ a*norms
       
                    indirect_effect := a*b
                    total_effect := c + (a*b)
                    '

greitemeyer_fit <- lavaan::sem(model = greitemeyer_mod,
                               data = greitemeyer_tib,
                               missing = "FIML",
                               estimator = "MLR")

model_performance(greitemeyer_fit) |> 
  display()

model_parameters(greitemeyer_fit) |> 
  display()

model_parameters(greitemeyer_fit, standardize = TRUE) |> 
  display()

```
