---
title: "An Adventure in Statistics"
subtitle: Getting Started with R, RStudio and Quarto
author: "Professor Andy Field"
institute: "University of Sussex"
format: 
  revealjs:
    auto-play-media: true
revealjs-plugins:
  - editable
filters:
  - editable
execute: 
  echo: true
  warning: false
  message: false
---

```{r}
#| echo: false

# general
library(easystats)
library(tidyverse)
# specific
library(emo)

source("../helpers/discovr_helpers.R")
source("../helpers/easystats_helpers.R")

set.seed(666)
cloak_tib <- discovr::invisibility_cloak |> 
  slice_sample(n = 3, by = cloak) |>  arrange(id)

rm_id <- discovr::invisibility_rm |>
  slice_sample(n = 5) |>  pull(id)

cloak_tib_rm <- discovr::invisibility_rm |> 
  filter(id %in% rm_id)

eddie_tib <- discovr::eddiefy
```


##  Key information

::: fragment

### Teaching

- 1 × weekly lecture (2 in weeks 1 and 2) on theory
- 1 × weekly practical class (2 in weeks 1 and 2) about `r rproj()`,  `r rstudio(scale = 0.3)` and `r quarto(scale = 0.3)`

:::
::: fragment

### Assessment

- 2 × 48-hour Take Away Papers (TAPs)
	- TAP 1: 25% (Week 6)
	- TAP 2: 30% (Week 9)
- Report: 45% (A1)

:::

::: notes
Use C to toggle pen/markup
Use backspace to delete markup
Use f to toggle fullscreen
:::

#  Part 1: Introducing `r rproj()` and `r rstudio(scale = 0.4)`

```{r, child = "why_r.qmd"}

```


#  Part 2: Workflow in `r rstudio(scale = 0.6)`

```{r, child = "workflow.qmd"}
```

#  Part 3: Introducing `r quarto(scale = 1)`


```{r, child = "quarto_intro.qmd"}
```


#  Part 4: Working with code 


```{r, child = "coding.qmd"}
```

#  Part 5: The `tidyverse` and `easystats`


```{r, child = "tidyverse.qmd"}
```

#  Part 6: Good habits

```{r, child = "tips.qmd"}
```


#  Part 7: Getting the most from practical classes

```{r, child = "practical_classes.qmd"}
```

